### 2024-02-19 outputs/paper/paper.qmd
finish graphs, edit some of the paper

---
title: "The World Bank, Women, Business and the Law"
subtitle: "The Need for better checking of Data"
author: 
  - Sehar Bajwa
  - Luca Carnegie
thanks: "Code and data are available at: https://github.com/lcarnegie/replicationpaper. "
date: February 20, 2024
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r workspace-setup}
#| include: false
#| warning: false
#| message: false

library(tidyverse)

## WBL Dataset Setup

wb_data <- read_csv(here::here("outputs/data/wb_data_clean.csv"))
```

# Introduction

# Data {#sec-data}

## Source {#subsec1-data}
## Methodology {#subsec2-data}

# Results {#sec-results}

# Discussion

## First discussion point {#sec-point1}

## Second discussion point {#sec-point2}

## Third discussion point {#sec-point3}

## Weaknesses and next steps

\newpage

# References

::: {#refs}
:::

\newpage

\appendix

# Appendix {-}